On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-01-18 MS.tex
Add new task to do on ms.tex
some changed in Material and methods and results

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- labels causing plot size to change, e.g. Fig 3d
%   -- only one y label for multi-panel plots
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis

@@

Compiled data included studies differing in the growth measurement used (RGR or AGR), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stage used to measure trait and growth rate. To assess whether such factors influenced our results, we investigated whether results changed between the entire dataset and a more conservative subset. The entire dataset includes all noted correlations. A second ``ideal'' subset, included only studies recording RGR across \underline{\textgreater} 10 species, under unstressed conditions, and where growth rates and traits were measured on plants from the same stage (expect for Hmax and seed mass, where a particular stage is by definition required to measure the trait).

@@

\subsection*{Trait coverage}
In total, 103 articles containing 551 correlations linking growth rate to any of the 19 traits  across a range of plant stages were identified (Fig. \ref{fig:fig1}). Considering only our top five traits of interest, the number reduced to 93 articles and 267 correlations. For measurements under ideal conditions, we identified only 55 studies with 137 correlations. The main results reported here were obtained with this ``ideal'' subset of data. These were then compare against results obtained with the entire dataset.

@@

Overall, results were consistent with predictions, although with limitations arising for some traits due to very unequal numbers of observations between growth stages.

@@

For WD, Hmax, and seed mass, correlations with RGR were also consistent with expectations, although our ability to test all hypotheses was at times limited due to low sample size. Indeed, CIs of $r$ often spanned zero especially where there were small numbers of studies for particular growth stages, precluding strong conclusions.
For WD, results were consistent with the prediction for negative correlations with RGR across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were not improved when stage was incorporated. Note that the CIs were not crossing zero for sapling and seedling stage despite the small sample sizes (n= 4 and 1 respectively), meaning a strong correlation and /or a good consistency in the coefficient reported for each stage across studies.

For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping), but also was consistent with no correlation (Fig \ref{fig:fig3}c, CIs overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives.

@@

\subsection*{Sensitivity of results to analysis decisions}

The above results were generated using data measured under ``ideal'' conditions: unstressed growth, growth measured as RGR, traits and growth recorded in same stage, \underline{\textgreater} 10 species. We now report results from the entire dataset, to determine if additional data would alter our conclusions. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species, and did not control for the source of trait or growth (i.e. both could be measured on different stages) or environmental conditions. These relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations. These additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had \textless five correlation coefficients: (i.e. WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

% TO DO : maybe explain in one sentence that mixing RGR and AGR is not a heresy here because we are not expecting different pattern with RGR vs AGR except for seedmass in seedling

Effect sizes and overall results using the entire dataset were similar to those reported above using the ``ideal'' subset (Fig. \ref{fig:fig4}). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across plant stage. The results were consistent with the prediction for a positive correlation with RGR across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect  between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with RGR at seedling stage; as the correlation was still positive  when 8 correlations measured under low light conditions were added.

For some trait-stage combinations, it was possible to compare results between RGR and AGR. For both WD and Hmax, effects size measured with RGR or AGR were no different at adult stage; CIs were overlapping despite equal sample size (Fig \ref{fig:fig4}b-c).  For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping).

@@

In total, we found only 103 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.

>>>>

% Todo:
% Ditch abbrvos for SLA, wood density, Hmax?
% Fix plots
%   -- major revision of code -- > lots of repetition, hard to follow
%   -- Figure 2: add * against traits used in meta-analysis
%   -- check the appendix: in particular test of heterogeneity etc. need to re-run the analysis
%   -- check the results and adequation with text of ms

@@

The entire dataset included studies differing in the growth measurement used (RGR or AGR), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stage used to measure trait and growth rate. The main results reported here were obtained with an ``ideal'' subset of data. This ``ideal'' subset, included only studies recording RGR across \underline{\textgreater} 10 species, under unstressed conditions, and where growth rates and traits were measured on plants from the same stage (expect for Hmax and seed mass, where a particular stage is by definition required to measure the trait). These were then compare against results obtained with the entire dataset, including all noted correlations.

@@

@@

Overall, results were consistent with predictions, although with limitations arising for some traits due to very unequal (and limited) numbers of observations between growth stages. We identified only 55 studies with 137 correlations reporting data measured under ``ideal'' conditions: unstressed growth, growth measured as RGR, traits and growth recorded in same stage, and correlation coefficient calculated across \underline{\textgreater} 10 species.

@@

For WD, Hmax, and seed mass, correlations with RGR were also consistent with expectations but our ability to test all hypotheses was at times limited due to low sample size. Indeed, while meta-analysis are possible even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e CI often spanned zero). 
For WD, results were consistent with the prediction for negative correlations with RGR across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginaly improved when stage was incorporated (p.value = 0.049). Note that the CIs were not crossing zero for seedling and sapling stage despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistency in the coefficient reported for each stage across studies.

For Hmax, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Here, Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not correlated at sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stages. 
While these results were consistent with expectation, we could not distinguish with another alternative: Hmax was correlated to growth for adult stage, but the correlation between Hmax and growth at sapling stage was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between Hmax and growth at small plant size.

@@

\subsection*{Main effects under relaxed criteria}

We now report results from the entire dataset, to determine if additional data would alter or reinforce our conclusions, drawn sometimes with limited replicates. This entire dataset included correlations measured with both AGR and RGR (Fig. \ref{fig:fig4}), across at least two species (but measured using several individuals), and did not control for the source of trait or growth (i.e. trait could be measured on a different stage) or environmental conditions. The relaxed criteria considerably increased the number of replicates, from $n = 137$ to $n = 267$ correlations across 93 articles. However, the additional replicates were unequally distributed across trait-stage combinations. Some trait-stage combinations already with a lot of data got even more support (e.g. from 25 to 59 correlation coefficients reported for SLA at seedling stage), while others still had few replicates. On the 13 traits-stage combinations studied in total, six had $\le$ five correlation coefficients: (i.e. WD-seedling, Hmax-sapling, Hmax-seedling, Seed mass-sapling, Aarea-adult and Aarea-sapling.

@@

Results from the entire dataset reinforced our conclusions for both SLA and WD, set appart alternative explanations for seed mass, and allowed to compare correlations with growth along plant stages for Aarea. For SLA and WD, the number of replicates doubled by using the entire dataset, and effect sizes and overall results were similar to those reported above using the ''ideal'' subset (Fig. \ref{fig:fig4})). As expected for seed mass, the correlation with growth at adult stage was significantly different from that observed at seedling stage (see CIs in Fig. \ref{fig:fig4}c, and model fit improved when stage was incorporated). For Aarea, the additional replicates in the entire dataset allowed us to compare the effect size across plant stage. The results were consistent with the prediction for a positive correlation with RGR across adult and seedling stages (Fig \ref{fig:fig4}e; CI not overlapping zero). Comparison of Aarea effect between the ``ideal'' subset (Fig. \ref{fig:fig3}d) and entire dataset (Fig. \ref{fig:fig4}d), indicated little effect of stress (here mainly low light) on the correlation with RGR at seedling stage; as the correlation was still positive when 8 correlations measured under low light conditions were added. For Hmax, there were still too few replicates to conclude undoubtely on the correlation to be absent for smaller sizes.

We had no hypotheses about a change in the correlation between trait and growth as AGR was used instead of RGR, except for seed mass. For seed mass, effects size differed significantly between RGR and AGR at seedling stage; as expected the correlation with RGR was positive and shifted to non significant with AGR (Fig \ref{fig:fig4}d; CI not overlapping). For other traits, effect size were generally consistent between RGR and AGR when both measurement were available. Some differences occured but they were not easily attributed to any single cause. For instance for WD, the correlation at juvenile stage shifted from negative with RGR to non significant with AGR (i.e CI included zero, Fig. \ref{fig:fig4}b). The correlation coefficients recorded for AGR came from Augspurger & Kelly (1984) and unlike others they used wood density values from adult plants together with growth from juveniles.

@@

In total, 103 articles published between 1983 and 2014, and containing 551 correlations linking growth rate to any of the 19 traits across a range of plant stages were identified (Fig. \ref{fig:fig1}). Climate zones spanned from temperate regions to tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. The number of species used for each correlation ranged from 2 to 300 species, with an average of 31. Given this variety, it was not surprising to find a wide range of $r$ reported for each trait. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}); suggesting confounding with these factors is not a likely explanation of the major results reported.
